---
title: "chapter3.Rmd"
author: "Yuwen Pang"
date: "09/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# chapter 3 Logistic regression

```{r}
date()
```

Here we go again...
In this chapter I practiced how to explore the raw data and build a Logistic regression; I have learnt and practiced several new functions and systematically understood the logistic regression.

- 1 for binary factor, i could use the logistic regression models, and these processed integrated glm()

- 2 before build the model, the data exploration seems quite important, which could help us to find the most potential parameters. in this case, I identified the potential correlation among high_use, sex, failures and other varietals.

- 3 the model validation: the perdition steps and calculated the probability, prediction, and then the error rate could be reports by user made function.
the k-folder validation would be more robust than one-time average error rate, and it also could be used to compare different models.



