# Insert chapter 2 Linear regression

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.

```{r}
date()
```

Here we go again...
In this chapter I practiced how to explore the raw data and build a linear regression;

1 load the data to local memory

2 explore the original data by selecting and filter

3 before build the regression, the pairs could help we to identify the potential explanatory variables

4 the model could be conduce perfectly at first time, we could improve it generally

5 the improving processes would use diagnostic plots and the summary of models

6 finally, we could get a improved and acceptable models, otherwise, the linear regression would not fit the case, we should look for nonlinear regressions, such as GAM.

# Insert chapter 3 Logistic regression

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.

```{r}
date()
```

Here we go again...