---
title: "chapter6"
author: "Yuwen Pang"
date: "23/11/2021"
output: html_document
---

# Chapter 6 Analysis of longitudinal data
```{r}
options(knitr.duplicate.label = "allow")
```

```{r}
date()
```

## 1 Load the longitudinal bprsl and ratsl datasets
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)

# Set the working directory
setwd("Z:/Course/Opendata science/IODS-project")

# read the longitudinal and original datasets
rats <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", sep="", header=TRUE)
bprs <- read.csv("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/BPRS.txt", sep="", header=TRUE)

bprsl <- read.csv('data/bprsl.csv', sep = ",", header = TRUE, stringsAsFactors = TRUE) 
ratsl <- read.csv('data/ratsl.csv', sep = ",", header = TRUE, stringsAsFactors = TRUE)

bprsl <- bprsl[,-1]
ratsl <- ratsl[,-1]

# Dimension of two form of data
dim(rats) #16*13
dim(ratsl) #176*5
dim(bprs) #40*11
dim(bprsl) #360*5

```

## 2.1 Summary Measure Analysis for ratsl data, the Chapter8 analysis
## Graphical Displays and Summary Measure Approach
Interpretations:
-1 in ratsl data, the weight ~ Time + Group + ID

-2 Base on graphs, we could see:
(1) wight increased with time in all groups

(2)the weight trends clearly separated in three groups:in which group1 had quite low values, and group 2 and 3 shard similar wight regions, from 400-550.

(3) Additionally, one ID in Group2 exceptional high weight, ranged from 550 to 640.

```{r message=FALSE, warning=FALSE}
names(ratsl)
str(ratsl)

# Convert Group and ID as factor varibles
ratsl$ID <- as.factor(ratsl$ID)
ratsl$Group <- as.factor(ratsl$Group)
ratsl$Time <- as.factor(ratsl$Time)
str(ratsl)

# Draw the plot of ratsl
# Plot1
ggplot(ratsl, aes(x = Time, y = weight, color = ID)) +
  geom_point(size=0.5) +
  geom_line(aes(group = ID)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(ratsl$weight), max(ratsl$weight)))


```

## 2.2 Summary Measure Analysis- Line plots
### the standarized plots and mean+sd plots
Interpretations:
-1 plot standardized weight, the differences between groups of weight trends with time were more apparent.

-2 In group1 and 3, there was one ID showed a significant low standardized weight trend line, while in group2, one curve/line posed in high value regions.

-3 here, i could more clearly found that the weight development with time varied in three groups, in which upward and downward trend both presented.

-4 the plot2 clearly shown the values gaps among three groups and illustrated their temporal trends, in which group 2 and 3 visible increased while group relatively stable.

```{r message=FALSE, warning=FALSE}
# Based on previous plots, I firstly group ratsl by Time
# Standardise the variable weight by Time
ratsl.gb <- ratsl %>%
  group_by(Group, Time) %>%
  mutate(stdweight = (weight - mean(weight))/sd(weight)) %>%
  ungroup()
dim(ratsl) #176*5
dim(ratsl.gb) #176*6
names(ratsl)

# Plot again with the standardised weight
ggplot(ratsl.gb, aes(x = Time, y = stdweight, linetype = ID)) +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  geom_line(aes(group = ID)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "standardized weight") +
  theme_bw()

# Number/levels of Times
n <- ratsl$Time %>% unique() %>% length()

# Summary rats with mean and standard error of weight by Group and Time 
ratss.gb <- ratsl %>%
  group_by(Group, Time) %>%
  summarise(mean = mean(weight), se = sd(weight)/sqrt(n)) %>%
  ungroup()
ratss.gb %>% summary()
dim(ratss.gb) #33*4

# Glimpse the grouped ratss data
glimpse(ratss.gb)

# Plot the mean profiles
names(ratss.gb) #"Group" "Time"  "mean"  "se"

ggplot(ratss.gb, aes(x = Time, y = mean, color = Group)) +
  geom_point(size=0.8) +
  geom_line(aes(group = Group)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width=0.3) +
  scale_y_continuous(name = "mean(weight) +/- se(weight)") +
  theme_bw()

```

## 2.3 Summary Measure Analysis- box plots
### found out the outliers
-1 by the box plot, i could see there were two outliers in group2 and 3, respectively.

-2 It's seems a bit diffecult to remove the outliers by each group, so i just removed the bigest outlier occured in Group 2

```{r message=FALSE, warning=FALSE}
# Create a summary data by GROUP and ID
ratss.gb2 <- ratsl %>%
  group_by(Group, ID) %>%
  summarise(mean=mean(weight)) %>%
  ungroup()
dim(ratss.gb2) #16*3
# Glimpse the data
glimpse(ratss.gb2)

# Draw a boxplot of the mean versus group
ggplot(ratss.gb2, aes(x = Group, y = mean, color = Group)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(weight)") +
  theme_bw()

# Remove the outliers
ratss.gb3 <- ratss.gb2 %>%
  filter((Group==1 & mean > 250)|(Group==2 & mean < 550)| (Group==3 & mean > 500))
dim(ratss.gb3) #15*3
dim(ratss.gb2) #16*3

# Draw the new boxplot of the mean versus group
ggplot(ratss.gb3, aes(x = Group, y = mean, color = Group)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(weight)") +
  ggtitle("the outlier filtered ratss") +
  theme_bw()

```

## 2.4 Summary Measure Analysis- fit the model and implied the t-test
-1 the linear model,fit1,  reported that the group variable is significant

-2 the fit2 revealed that time and group both significantly impact weight
```{r message=FALSE, warning=FALSE}
# Fit the linear model with the mean as the response 
fit1 <- lm(mean ~ Group, data = ratss.gb3)

# Compute the analysis of variance table for the fitted model with anova()
anova(fit1)

# Fit the linear model with the mean as the response
names(ratss.gb)
# "Group" "Time"  "ID"    "mean"  "se"  

fit2 <- lm(mean ~ Group + Time, data = ratss.gb)

# Compute the analysis of variance table for the fitted model with anova()
anova(fit2)
```


## 3 Linear Mixed Effects Models for bprsl data, the Chapter9 analysis
-1 bprs ~ treatment + subject + week, the potential formulation

-2 the week would be the random effect
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)

# Load bprs and bprsl
bprs <- read.csv("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/BPRS.txt", sep="", header=TRUE)
bprsl <- read.csv('data/bprsl.csv', sep = ",", header = TRUE, stringsAsFactors = TRUE) 
bprsl <- bprsl[,-1]

# the dimension of two form of data
dim(bprs) #40*11
dim(bprsl) #360*6

names(bprsl)
str(bprsl)

# Convert the categorical variables
bprsl$treatment <- as.factor(bprsl$treatment)
bprsl$subject <- as.factor(bprsl$subject)

# Draw the plot of new long data form
ggplot(bprsl, aes(x = week, y = bprs, group = subject)) +
  geom_text(aes(label = treatment, color = treatment)) +
  # facet_grid(. ~ treatment, labeller = label_both) +
  scale_x_continuous(name = "week", breaks = seq(0, 8, 1)) +
  scale_y_continuous(name = "bprs", limits = c(min(bprsl$bprs), max(bprsl$bprs))) +
  ggtitle("the orginal plots of bprsl") +
  theme_bw()

ggplot(bprsl, aes(x = week, y = bprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  scale_x_continuous(name = "week", breaks = seq(0, 8, 1)) +
  scale_y_continuous(name = "bprs", limits = c(min(bprsl$bprs), max(bprsl$bprs))) +
  ggtitle("the orginal line plots of bprsl") +
  theme_bw()

```

## 3.2 Exploring bprsl
-1 grouped bprs by subjects, and saw how each treatment varied with week/time

-2 from this line plot, i could clearly find that the impact of teatment and week on bprs, while these two varibles further affecte by subject that acts as random effect or a indepedent varible.
```{r}

# Number of subject
n <- bprsl$subject %>% unique() %>% length()

# Summary data with mean and standard error of bprs by treatment and week 
bprss <- bprsl %>%
  group_by(treatment, week) %>%
  summarise( mean = mean(bprs), se = sd(bprs)/sqrt(n) ) %>%
  ungroup()

# Plot the mean profiles of subjects
ggplot(bprss, aes(x = week, y = mean, linetype = treatment, shape = treatment)) +
  geom_line(aes(group = treatment)) +
  scale_linetype_manual(values = c(1,2)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, linetype="1"), width=0.3) +
  scale_y_continuous(name = "mean(bprs) +/- se(bprs)") +
  theme_bw()

```


## 3.3 fit and compare models
-1 the independent variable (or random effect) was 'subject', and it influenced by week

-2 built two linear mixed model: fit1 and fit2

```{r}
# access library lme4
library(lme4)

# Create a random intercept model
bprsl_ref <- lmer(bprs ~ week + treatment  + (1 | subject), data = bprsl, REML = FALSE)

# Print the summary of the model
summary(bprsl_ref)

# create a random intercept and random slope model
bprsl_ref1 <- lmer(bprs ~ week + treatment + (week | subject), data = bprsl, REML = FALSE)

# print a summary of the model
summary(bprsl_ref1)

# perform an ANOVA test on the two models
anova(bprsl_ref1, bprsl_ref)


```

## 3.4 plot the fitted model
-1 based on the t-test, anova, the (week|subject) model was be accepted

-2 calculated the fitted values and draw the predicted plot
```{r message=FALSE, warning=FALSE}
# Create a vector of the fitted values
Fitted <- fitted(bprsl_ref1)

# Create a new column fitted to RATSL
bprsl <- bprsl %>%
  mutate(Fitted)

ggplot(bprsl, aes(x = week, y = Fitted, linetype = subject)) +
  geom_line(aes(group = subject)) +
  scale_linetype_manual(values = rep(1:10, times=2)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme_bw()

```

