# chapter 2 Linear regression

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.

```{r}
date()
```

Here we go again...
In this chapter I practiced how to explore the raw data and build a linear regression;
(Creat lists with a dash)

- 1 load the data to local memory


- 2 explore the original data by selecting and filter
(With R markdown, you can insert R code chunks including table and plots:)

```{r ggplot, fig.width=10, fig.hight=4, message=FALSE}
# quick summary of sudent2014 data
# library(dplyr)
library(ggplot2)
library(GGally)
# Load the student2014 data set
stu14 <- read.csv('http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/learning2014.txt')
# Summary the student2014
summary(stu14)
# Print the paris plots
p <- ggpairs(stu14, mapping = aes(col = gender,alpha = 0.3), 
         lower = list(combo = wrap("facethist", bins = 20)))
p
```


- 3 before build the regression, the pairs could help we to identify the potential explanatory variables
```{r}
# As the pair correlation showed,
# the points as respond, the attitude, stra and surf as the explantory variables
model <- lm(points ~ attitude + stra + surf , data = stu14)
summary(model)
```

- 4 the model could be conduce perfectly at first time, we could improve it generally
```{r}
# Examine the model
par(mfrow = c(2,2))
plot(model, which = c(1,2,5))
```

- 5 the improving processes would use diagnostic plots and the summary of models
```{r}
# rebuild the fitted model
stu14_2 <- stu14[-c(35,56,77,145),] #162*7
model2 <- lm(points ~ attitude + stra , data = stu14_2)
summary(model2)
```

- 6 finally, we could get a improved and acceptable models, otherwise, the linear regression would not fit the case, we should look for nonlinear regressions, such as GLM.
```{r}
# the r2 has been improved a bit in model2, from 0.21 to 0.29
# the diagnostic plots for model 2
par(mfrow = c(2,2))
plot(model2, which = c(1,2,5))
```

